---
title: "Project 1"
author: "Cassie Boylan"
date: "9/19/2020"
output:

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
install.packages("data.table")
```

Loading libraries to use
```{r}
library(tidyverse)
library(ggplot2)
library(RCurl)
library(plyr)
library(dplyr)
library(stringr)
library(readr)
library(data.table)
```

## list of actions for cleaning

1. Read in text file
    use fill to handle jagged row error
 
 2. Name data.table columns
 3. Drop unused header data
 4. Drop rows with junk data

 5. Subset rows with number and name
 6. Subset rows with state and pre-score values
 7. Create new column to hold Pre-Score values
      mutate with string exp?

 8. Extract the Contestant ID from each Round
   
 9. Match each Round contestant's pair ID to contestant and look at contestant pre-score
   
 10. Sum contestant pre-score and divide by # of contestants to calculate Average Pre-chess Rating'
  
 11. Create new column to hold average pre-chess rating
 
```{r read-in-data}

#x <- getURL('https://raw.githubusercontent.com/cassie-boylan/DATA-607-Project-1/master/tournamentinfo.txt')

tournament <- read.table('https://raw.githubusercontent.com/cassie-boylan/DATA-607-Project-1/master/tournamentinfo.txt', 
          stringsAsFactors = FALSE,
           sep = "|",
          # colClasses = c("factor",rep("character",times =10)),
           nrows = -1,
           skip = 1,
           fill = TRUE,
           strip.white = TRUE,
           comment.char = "",
           allowEscapes = TRUE)

```

Set Column Names
```{r set-column-names}
names(tournament)[1] <- "State"
names(tournament)[2] <- "Name"
names(tournament)[3] <- "Total Points" 
names(tournament)[4:10] <- paste("Round", 1:7)
```

Reshape Data Table
```{r drop-duplicate-header & extra-column}
tournament_clean <- tournament[-c(1:3), -11]
```

Reindex Data Table
```{r reindex-data.table}
rownames(tournament_clean) <- 1:nrow(tournament_clean)
```

Subset into Multiple Data Tables to filter & extract text values
```{r pull out pair-number-and-name}
contestants <- tournament_clean %>% 
  select(Pair_ID = State, Name, `Total Points`) %>%
  filter(Pair_ID %in% c(grep("^[0-9]+",Pair_ID, value=TRUE)))
```

```{r pull out state-and-prescore}
contestants_2 <- tournament_clean %>% 
  select(State, Pre_Score_Raw = Name) %>%
  filter(State %in% c(grep("^[A-Z]+",State, value=TRUE)))
```


```{r extract-score-with-tmp-columns}
contestants_2.0 <-contestants_2 %>%
  mutate(score = str_sub(Pre_Score_Raw, 14)) %>%
  mutate(score2 = str_extract(score,"[0-9]{3,4}"))

```

```{r drop-tmp-columns}

contestants_2.1 <- contestants_2.0 %>%
  select(State, Pre_Score = score2)
```

```{r}
rounds <- tournament_clean %>%
  filter(`Round 1` %in% c(grep("[0-9]",`Round 1`, value=TRUE))) %>%
  transmute(Pair_ID = State, 
              Contestant_1 = str_sub(`Round 1`,-2),
              Contestant_2 = str_sub(`Round 2`,-2),
              Contestant_3 = str_sub(`Round 3`,-2),
              Contestant_4 = str_sub(`Round 4`,-2),
              Contestant_5 = str_sub(`Round 5`,-2),
              Contestant_6 = str_sub(`Round 6`,-2),
              Contestant_7 = str_sub(`Round 7`,-2))
```

Merge Separate Data Tables Together
```{r merge-clean-data}
new_df <- merge(contestants, contestants_2.1, by=0, all=TRUE)
```

```{r}
total_df <- merge(new_df, rounds, by="Pair_ID", all=TRUE)
```

Summarize All Contestants Pre-Scores & Divide by Matches
<!--where I ran out of time
{r}
total_df %>%
  summarize(
    contestant_summary_score = )
 ->
```

